---
title: "Incarceration Data"
author: "Samiira Yusuf"
date: '2022-11-01'
output: html_document
---

```{r echo=FALSE, eval=TRUE, message=FALSE}
source("../source/a4-helpers.R")
chooseCRANmirror(graphics = FALSE, ind = 1)
source("../source/analysis.R")
```

## Abstract 
One of the biggest problems in America is the high rates of incarceration. The incarceration rate in the United States is a major issue. The vast majority of the millions of prisoners are members of various minorities. Police frequently target Black and Latinx people, which explains why there is so much injustice in the court system. This information demonstrates how inequitable incarceration is in our nation.



## Introduction 
The topic of this project is incarceration in each state of the United States. It includes several elements, such as various racial and gender groups. America has a serious problem with incarceration. Numerous minorities make up a large portion of the millions of inmates. Black people and Latinx people are overpolicing and being targeted by police, which is why there is so much inequity in the court system. The incarceration trends data set includes data from 1970 to 2018; over that time, there has been a marked increase, and as each year goes by, the problem only becomes worse in America. People who are incarcerated now or have been in the past are the direct stakeholders who are influenced by this dataset research since they are included in it and are the ones who are. The relatives of those who are or were incarcerated and minorities are indirect stakeholders. I say this because policies frequently target minorities, and the jail system has the potential to have a detrimental impact on minorities at any time.

## Data Summary 
 The 50 states included in this dataset each have a sizable amount of incarceration data. `r max_black_jail_state` is the state where there are the most black inmates. In the most recent year, there has been `r black_max_pop` black people in jail. In addition, there has been `r latinx_max_pop` Latinx people in jail in the most recent year. On the other hand, there has been `r white_max_pop`r white people in jail the same year, which is a significant difference difference from the black and latinx population
  
## Section 3
```{r echo = FALSE}
plot_jail_pop_for_us()
```


*Caption*:Data distribution of the inmate population from 1970 to 2018.


*Summary*: This bar graph compares the overall inmate population for each year. The earliest years have a smaller population than the latter years, as shown in this graph. The population began growing gradually in 1980, and by 1990, it had significantly increased. Additionally, the population peaked between 2006 and 2009. Why did the population grow so much after 1990 is a question I have.


## Section 4
```{r echo = FALSE}
results_line_plot
```



*Caption*: Jail population for California, Minnesota, Florida, and Washington.


*Summary*: This line chart compares jail populations for four states, which are California, Minnesota, Florida, and Washington. As seen in this line chart, California has the highest jail population, starting off pretty high as well. Florida starts off pretty low but then increases significantly as the years go on. However, Minnesota and Washington's jail population is much less than Texas and California's. Their population never passes 20,000. A question I have is, why did Florida’s jail population has increased so fast?


## Section 5
```{r echo = FALSE}
compare
```


*Caption*: Comparing black and Latinx jail populations in 2009 to 2012.

*Summary*: I chose to compare the jail population for these two ethnic groups, I looked at the years 2009 to 2012. This bar graph shows that the Latinx population has increased over the course of four years, notably in 2009. Their populations actually differ very much, which surprises me because I didn't expect the white jail population to be more. A question I have is what year was the population the highest for both ethnic groups? And by looking at this bar chart, it is clear that the year it was highest for Latinx people was 2009. It was slightly highest for black people in 2009 as well.


## Section 6
```{r echo = FALSE}
black_jail
```



*Caption*: Black jail population in different areas of Texas. 

*Summary*: This map shows the jail population in each county in Texas. Most of the counties have a lighter shade which shows that there isn't a very high population in that area. However, there are three counties on the map that do have a dark shade of blue which shows that the jail population is really high there. A question I have is why is that one area really dark compared to the rest.
